Estimation of Intracortical Connectivity in a Dynamical Neural Field Model

### 2010-02-11 ltx/BrainIDE.tex
some things

The average firing rate of the presynaptic neurons is related to the postsynaptic membrane potential by the sigmoidal activation function
\begin{equation}\label{ActivationFunction}
f\left( {{v_t}\left( {r'} \right)} \right) = \frac{{{\nu _0}}}{{1 + \exp \left( {\beta \left( {{v_0}\left( r \right) - {v}\left(r',\bar{t} \right)} \right)} \right)}}.
\end{equation}
%\begin{equation}\label{RateOfPreSynPops}
%p\left( {v\left( {r,\bar t} \right)} \right) = {\nu_0}u\left( {v\left( {r,\bar t} \right) - {v_0}\left( r \right)} \right)
%\end{equation}
%\begin{equation}\label{HeavisideDef}
%u\left( {v\left( {r,\bar t} \right) - {v_0}\left( r \right)} \right) = \left\{ {\begin{array}{*{20}{c}}
%   1 & {v\left( {r,\bar t} \right) \ge {v_0}\left( r \right)}  \\
%   0 & {otherwise}  \\
%\end{array}} \right.
%\end{equation}
Even though we will model  $v_0$ as being homogeneous across the cortical sheet we will express it as being spatially dependant allowing for generality when constructing the state-space representation in the next section. The parameters in equation~\ref{ActivationFunction} are detailed in Table 1. By combining equations~\ref{SpikesToPotential} and \ref{RateBasedInteractions} we get the spatiotemporal model
\begin{equation}\label{FullDoubleIntModel}
v\left(r,t\right) = \int\limits_{-\infty}^t\int\limits_\Omega  h\left(t - t'\right)w\left(r,r'\right)f\left( v\left( r',\bar t \right)\right)dr' dt'+e\left(r,t\right)
\end{equation}
To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function
\begin{equation}\label{GreensFuncDef}
Lh\left( t \right) = \delta \left( t \right),
\end{equation}
where $L$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function. This provides the most general form
\begin{equation}\label{GenForm}
L\left( {\frac{\partial }{{dt}}} \right)v\left( {r,t} \right) = \int_\Omega  {w\left( {r,r'} \right)f\left( {v\left( {r',\bar t} \right)} \right)dr'}  + e\left( {r,t} \right),
\end{equation}
where $L$ is a polynomial of order $n$ with constant coefficients that provides an $n^{th}$ order model. Most studies considering this model use first or second order time derivatives. We shall use the first order model giving
\begin{equation}\label{FinalForm1}
\frac{{dv\left( {r,t} \right)}}{{dt}} + \alpha{v\left( {r,t} \right)} = \int_\Omega  {w\left( {r,r'} \right)f\left( {v\left( {r',\bar t} \right)} \right)dr'}  +e\left( {r,t} \right).
\end{equation}
\subsection{Multi-Layer Model}
\begin{equation}\label{ExcitatoryLayer}
\frac{dv_e\left( {r,t} \right)}{{dt}} + \alpha_ev_e\left( {r,t} \right) = \int_\Omega  {w_{ee}\left( {r,r'} \right)f\left( v_e\left( r',\bar t \right) \right) + w_{ei}\left( {r,r'} \right)f\left( v_i\left( r',\bar t \right) \right)dr'}  +\varepsilon_e\left( {r,t} \right)
\end{equation}
\begin{equation}\label{InhibitoryLayer}
\frac{dv_i\left( {r,t} \right)}{{dt}} + \alpha_iv_i\left( {r,t} \right) = \int_\Omega  {w_{ie}\left( {r,r'} \right)f\left( v_e\left( r',\bar t \right) \right) + w_{ii}\left( {r,r'} \right)f\left( v_i\left( r',\bar t \right) \right)dr'}  +\varepsilon_i\left( {r,t} \right)
\end{equation}
The subscript $ei$ denotes connectivity from an inhibitory neural element to an excitatory neural element. We can write the model in matrix form
\begin{equation}\label{matrixEquation}
\left[ {\begin{array}{*{20}{c}}
   {\frac{{d{v_e}\left( t \right)}}
{{dt}}}  \\
   {\frac{{d{v_i}\left( t \right)}}
{{dt}}}  \\

 \end{array} } \right] = \int_\Omega  {\left[ {\begin{array}{*{20}{c}}
   {{w_{ee}}\left( {r,r'} \right)} & {{w_{ei}}\left( {r,r'} \right)}  \\
   {{w_{ie}}\left( {r,r'} \right)} & {{w_{ii}}\left( {r,r'} \right)}  \\

 \end{array} } \right]} \left[ {\begin{array}{*{20}{c}}
   {f\left( {{v_e}\left( {r',\bar{t}} \right)} \right)}  \\
   {f\left( {{v_i}\left( {r',\bar{t}} \right)} \right)}  \\

 \end{array} } \right]dr' - \left[ {\begin{array}{*{20}{c}}
   {{\alpha _e}{v_e}\left( {r,t} \right)}  \\
   {{\alpha _i}{v_i}\left( {r,t} \right)}  \\

 \end{array} } \right] + \left[ {\begin{array}{*{20}{c}}
   {{\varepsilon _e}\left( t \right)}  \\
   {{\varepsilon _i}\left( t \right)}  \\

 \end{array} } \right]
\end{equation}

\begin{equation}\label{matrix_equation_in_vector_notation}
   \frac{d\mathbf{V}(r,t)}{dt} = \int_\Omega \mathbf{W}(r,r')f\left(\mathbf{V}(r,t)\right)dr' - \mathbf{A}\mathbf{V}(r,t) + \mathbf{E}(r,t)
\end{equation}

The iEEG observation equation is
\begin{equation}\label{iEEGObservationEquation}
    y(k,t) = \int_{\Omega} m(k,r)\left(v_e(r,t) + v_i(r,t)\right)dr' + w(k,t)
\end{equation}

\subsection{Alternate Model Description I}
An alternate and equivalent way to remove the temporal integration using the Laplace transform . We start by considering a discrete neural mass $i$ with inputs with zero delay from another discrete mass $j$ and
\begin{equation}\label{TwoDiscreteMasses}
{v_i}\left( t \right) = \int_{ - \infty }^t {{h_j}(t - t'){w_{ij}}p\left( {{v_j}\left( t \right)} \right)dt'},
\end{equation}
where $w_{ij}$ can be considered the synaptic efficiency from $j$ to $i$. Now we take the Laplace transform $\mathcal{L}(v_i(t))=V(s)$ yielding
\begin{eqnarray}
% \nonumber to remove numbering (before each equation)
  {V_i}\left( s \right) &=& {H_j}\left( s \right){w_{ij}}P\left( {{V_j}\left( s \right)} \right) \nonumber \\
   &=& {\left( {s + \alpha} \right)^{ - 1}}{w_{ij}}P\left( {{V_j}\left( s \right)} \right)  \\
  \Rightarrow s{V_i}\left( s \right) + \alpha{V_i}\left( s \right) &=& {w_{ij}}P\left( {{V_j}\left( s \right)} \right).
\end{eqnarray}
Now taking the inverse Laplace transform we get
\begin{equation}\label{InverseLaplace}
\frac{{d{v_i}\left( t \right)}}{{dt}} + \alpha{v_i}\left( t \right) = {w_{ij}}p\left( {{v_j}\left( t \right)} \right).
\end{equation}
Next we convert the model from discrete neural masses to a continuum by
\begin{equation}\label{Discrete2Continuum}
\frac{{dv\left( {r,t} \right)}}{{dt}} + \alpha{v\left( {r,t} \right)} = \int_\Omega  {w\left( {r,r'} \right)p\left( {v\left( {r',\bar t} \right)} \right)dr'}  + e\left( {s,t} \right).
\end{equation}

@@

\section{State-Space Model}
We begin the derivation of the state-space model with the discrete time, continuous space IDE neural field model. A one-step Euler method was used to transform the ODE to a discrete time form.
\begin{equation}\label{SystemEq}
\frac{v_{t + 1}\left( \mathbf{r} \right) - v_t(\mathbf{r})}{T_s} = \int_\Omega  {k\left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{v_t}\left( {{\bf{r'}}} \right)} \right)d{\bf{r'}}}  - \alpha {v_t}\left( {\bf{r}} \right) + e_t\left( {\mathbf{r}} \right),
\end{equation}
\begin{equation}\label{SystemEq}
v_{t + 1}\left( \mathbf{r} \right)  = T_s\int_\Omega  {k\left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{v_t}\left( {{\bf{r'}}} \right)} \right)d{\bf{r'}}} - T_s\alpha {v_t}\left( {\bf{r}} \right) + v_t(\mathbf{r}) + T_se_t\left( {\mathbf{r}} \right),
\end{equation}
\begin{equation}\label{SystemEq}
v_{t + 1}\left( \mathbf{r} \right)  = T_s\int_\Omega  {k\left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{v_t}\left( {{\bf{r'}}} \right)} \right)d{\bf{r'}}} + (1 - T_s\alpha) v_t\left( {\bf{r}} \right) + T_se_t\left( {\mathbf{r}} \right),
\end{equation}
\begin{equation}\label{SystemEq}
v_{t + 1}\left( \mathbf{r} \right)  = T_s\int_\Omega  {k\left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{v_t}\left( {{\bf{r'}}} \right)} \right)d{\bf{r'}}} + \lambda v_t\left( {\bf{r}} \right) + T_se_t\left( {\mathbf{r}} \right).
\end{equation}
where $e_t(r)\sim \mathcal{GP}(0,R)$, $\lambda = 1-T_s\alpha$ and $T_s$ is the time step or sampling step.

@@

In order to implement standard estimation techniques we propose a decomposition of the The field using a set of Gaussian basis functions. Decomposition allows a continuous field to be represented by a finite dimensional state vector. This allows for the application of standard nonlinear, non-Gaussian state estimation methods such as sequential Monte-Carlo techniques like particle filtering. The field decomposition is described by
\begin{equation}\label{DefFieldDecomp}
{v_t}\left( {\bf{r}} \right) = {\phi ^T}\left( {\bf{r}} \right){x_t}.
\end{equation}
where $\phi(r)$ is the Gaussian basis functions that are scaled by the the state vector. The connectivity kernel can all be decomposed into a set of basis functions
\begin{equation}\label{DefKernelDecomp}
k( \mathbf{r} , \mathbf{r'} ) = \theta^T\psi(\mathbf{r} , \mathbf{r'} )
\end{equation}
Will will assume we know the parametric form of the basis functions, where the parameter $\theta$ is unknown. Each connectivity basis function can be considered a layer in the WC model, representing short range excitation, surround inhibition and long range excitation. Making substitutions we get
\begin{equation}\label{DecompModel}
{\phi ^T}\left( \mathbf{r} \right){x_{t+1}} = \theta^T\int_\Omega  {\psi \left( {{\bf{r}} , {\mathbf{r'}}} \right)f\left( {{\phi ^T}\left( {{\mathbf{r'}}} \right){x_t}} \right)d{\mathbf{r'}}}  - \alpha\phi^T(\mathbf{r})x_t + {\varepsilon _t}\left( {\mathbf{r}} \right).
\end{equation}
To isolate the state we cross multiply by $\phi(\mathbf{r})$ and integrate over $\mathbf{r}$
\begin{equation}\label{Poo}
\int_\Omega  {\phi \left( {\bf{r}} \right){\phi ^T}\left( {\bf{r}} \right)d{\bf{r}}{x_{t+1}}}  = \int_\Omega  {\phi ( \mathbf{r} )\theta^T\int_\Omega  {\psi \left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{\phi ^T}\left( {{\bf{r'}}} \right){x_t}} \right)d{\bf{r'}}} d{\bf{r}}}  - \alpha\int_\Omega \phi(\mathbf{r})\phi^T(\mathbf{r})d\mathbf{r} x_t  + \int_\Omega  {\phi \left( {\mathbf{r}} \right){\varepsilon _t}\left( {\mathbf{r}} \right)d{\mathbf{r}}}.
\end{equation}
Now we define
\begin{equation}\label{DefGamma}
\Gamma  = \int_\Omega  {\phi \left( {\bf{r}} \right){\phi ^T}\left( {\bf{r}} \right)d{\bf{r}}}.
\end{equation}
Substituting equation~\ref{DefGamma} into \ref{Poo} and rearranging gives
\begin{equation}\label{SS Model}
{x_{t+1}} = {\Gamma ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right)\theta^T\int_\Omega  {\psi \left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{\phi ^T}\left( {{\bf{r'}}} \right){x_t}} \right)d{\bf{r'}}} d{\bf{r}}}-\alpha x_t + {\Gamma ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right){\varepsilon _t}\left( {\bf{r}} \right)d{\bf{r}}}.
\end{equation}
We can rearrange to get
\begin{equation}\label{Homogeneous SS Model}
{x_{t+1}} = {\Gamma ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right)\int_\Omega  {f\left( {{\phi ^T}\left( {{\bf{r'}}} \right){x_t}} \right)\psi^T \left( {{\bf{r}} - {\bf{r'}}} \right)d{\bf{r'}}} d{\bf{r}}}\theta - \alpha x_t + {\Gamma ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right){\varepsilon _t}\left( {\bf{r}} \right)d{\bf{r}}}.
\end{equation}
This provides the final state-space model
\begin{equation}\label{AbbrevSSModel}
x_{t + 1} = \theta^T q(x) -\alpha x + \Gamma ^{ - 1}\int_\Omega  {\phi ( \mathbf{r} ){\varepsilon _t}( \mathbf{r} )d\mathbf{r}}
\end{equation}
with observation equation
\begin{equation}\label{ObservationEquation}
    y_t( k ) = \int_\Omega  m( k,\mathbf{r} )\phi^T( \mathbf{r} ) d\mathbf{r}x_t  + \varepsilon ( k ),
\end{equation}

\section{Estimation of the Nonlinear Homogeneous IDE Neural Field Model}

>>>>

The average firing rate of the presynaptic neurons is related to the postsynaptic membrane potential by the sigmoidal activation function
\begin{equation}\label{ActivationFunction}
f\left( {{v_t}\left( {r'} \right)} \right) = \frac{{{\nu _0}}}{{1 + \exp \left( {\beta \left( {{v_0}\left( r \right) - {v}\left(r',\bar{t} \right)} \right)} \right)}}.
\end{equation}
%\begin{equation}\label{RateOfPreSynPops}
%p\left( {v\left( {r,\bar t} \right)} \right) = {\nu_0}u\left( {v\left( {r,\bar t} \right) - {v_0}\left( r \right)} \right)
%\end{equation}
%\begin{equation}\label{HeavisideDef}
%u\left( {v\left( {r,\bar t} \right) - {v_0}\left( r \right)} \right) = \left\{ {\begin{array}{*{20}{c}}
%   1 & {v\left( {r,\bar t} \right) \ge {v_0}\left( r \right)}  \\
%   0 & {otherwise}  \\
%\end{array}} \right.
%\end{equation}
Even though we will model  $v_0$ as being homogeneous across the cortical sheet we will express it as being spatially dependant allowing for generality when constructing the state-space representation in the next section. The parameters in equation~\ref{ActivationFunction} are detailed in Table 1. By combining equations~\ref{SpikesToPotential} and \ref{RateBasedInteractions} we get the spatiotemporal model
\begin{equation}\label{FullDoubleIntModel}
v\left(r,t\right) = \int\limits_{-\infty}^t\int\limits_\Omega  h\left(t - t'\right)w\left(r,r'\right)f\left( v\left( r',\bar t \right)\right)dr' dt'+e\left(r,t\right)
\end{equation}
To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function
\begin{equation}\label{GreensFuncDef}
Lh\left( t \right) = \delta \left( t \right),
\end{equation}
where $L$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function. This provides the most general form
\begin{equation}\label{GenForm}
L\left( {\frac{\partial }{{dt}}} \right)v\left( {r,t} \right) = \int_\Omega  {w\left( {r,r'} \right)f\left( {v\left( {r',\bar t} \right)} \right)dr'}  + e\left( {r,t} \right),
\end{equation}
where $L$ is a polynomial of order $n$ with constant coefficients that provides an $n^{th}$ order model. Most studies considering this model use first or second order time derivatives. We shall use the first order model giving
\begin{equation}\label{FinalForm1}
\frac{{dv\left( {r,t} \right)}}{{dt}} + \alpha{v\left( {r,t} \right)} = \int_\Omega  {w\left( {r,r'} \right)f\left( {v\left( {r',\bar t} \right)} \right)dr'}  +e\left( {r,t} \right).
\end{equation}

% \subsection{Multi-Layer Model}
% \begin{equation}\label{ExcitatoryLayer}
% \frac{dv_e\left( {r,t} \right)}{{dt}} + \alpha_ev_e\left( {r,t} \right) = \int_\Omega  {w_{ee}\left( {r,r'} \right)f\left( v_e\left( r',\bar t \right) \right) + w_{ei}\left( {r,r'} \right)f\left( v_i\left( r',\bar t \right) \right)dr'}  +\varepsilon_e\left( {r,t} \right)
% \end{equation}
% \begin{equation}\label{InhibitoryLayer}
% \frac{dv_i\left( {r,t} \right)}{{dt}} + \alpha_iv_i\left( {r,t} \right) = \int_\Omega  {w_{ie}\left( {r,r'} \right)f\left( v_e\left( r',\bar t \right) \right) + w_{ii}\left( {r,r'} \right)f\left( v_i\left( r',\bar t \right) \right)dr'}  +\varepsilon_i\left( {r,t} \right)
% \end{equation}
% The subscript  denotes connectivity from an inhibitory neural element to an excitatory neural element. We can write the model in matrix form
% \begin{equation}\label{matrixEquation}
% \left[ {\begin{array}{*{20}{c}}
%    {\frac{{d{v_e}\left( t \right)}}
% {{dt}}}  \
%    {\frac{{d{v_i}\left( t \right)}}
% {{dt}}}  \
% 
%  \end{array} } \right] = \int_\Omega  {\left[ {\begin{array}{*{20}{c}}
%    {{w_{ee}}\left( {r,r'} \right)} & {{w_{ei}}\left( {r,r'} \right)}  \
%    {{w_{ie}}\left( {r,r'} \right)} & {{w_{ii}}\left( {r,r'} \right)}  \
% 
%  \end{array} } \right]} \left[ {\begin{array}{*{20}{c}}
%    {f\left( {{v_e}\left( {r',\bar{t}} \right)} \right)}  \
%    {f\left( {{v_i}\left( {r',\bar{t}} \right)} \right)}  \
% 
%  \end{array} } \right]dr' - \left[ {\begin{array}{*{20}{c}}
%    {{\alpha _e}{v_e}\left( {r,t} \right)}  \
%    {{\alpha _i}{v_i}\left( {r,t} \right)}  \
% 
%  \end{array} } \right] + \left[ {\begin{array}{*{20}{c}}
%    {{\varepsilon _e}\left( t \right)}  \
%    {{\varepsilon _i}\left( t \right)}  \
% 
%  \end{array} } \right]
% \end{equation}
% 
% \begin{equation}\label{matrix_equation_in_vector_notation}
%    \frac{d\mathbf{V}(r,t)}{dt} = \int_\Omega \mathbf{W}(r,r')f\left(\mathbf{V}(r,t)\right)dr' - \mathbf{A}\mathbf{V}(r,t) + \mathbf{E}(r,t)
% \end{equation}
% 
% The iEEG observation equation is
% \begin{equation}\label{iEEGObservationEquation}
%     y(k,t) = \int_{\Omega} m(k,r)\left(v_e(r,t) + v_i(r,t)\right)dr' + w(k,t)
% \end{equation}{}

% \subsection{Alternate Model Description I}
% An alternate and equivalent way to remove the temporal integration using the Laplace transform . We start by considering a discrete neural mass $i$ with inputs with zero delay from another discrete mass $j$ and
% \begin{equation}\label{TwoDiscreteMasses}
% {v_i}\left( t \right) = \int_{ - \infty }^t {{h_j}(t - t'){w_{ij}}p\left( {{v_j}\left( t \right)} \right)dt'},
% \end{equation}
% where $w_{ij}$ can be considered the synaptic efficiency from $j$ to $i$. Now we take the Laplace transform $\mathcal{L}(v_i(t))=V(s)$ yielding
% \begin{eqnarray}
% % \nonumber to remove numbering (before each equation)
%   {V_i}\left( s \right) &=& {H_j}\left( s \right){w_{ij}}P\left( {{V_j}\left( s \right)} \right) \nonumber \\
%    &=& {\left( {s + \alpha} \right)^{ - 1}}{w_{ij}}P\left( {{V_j}\left( s \right)} \right)  \\
%   \Rightarrow s{V_i}\left( s \right) + \alpha{V_i}\left( s \right) &=& {w_{ij}}P\left( {{V_j}\left( s \right)} \right).
% \end{eqnarray}
% Now taking the inverse Laplace transform we get
% \begin{equation}\label{InverseLaplace}
% \frac{{d{v_i}\left( t \right)}}{{dt}} + \alpha{v_i}\left( t \right) = {w_{ij}}p\left( {{v_j}\left( t \right)} \right).
% \end{equation}
% Next we convert the model from discrete neural masses to a continuum by
% \begin{equation}\label{Discrete2Continuum}
% \frac{{dv\left( {r,t} \right)}}{{dt}} + \alpha{v\left( {r,t} \right)} = \int_\Omega  {w\left( {r,r'} \right)p\left( {v\left( {r',\bar t} \right)} \right)dr'}  + e\left( {s,t} \right).
% \end{equation}

@@

\section{State-Space Model}
We begin the derivation of the state-space model with the discrete time, continuous space IDE neural field model. A one-step Euler method was used to transform the ODE to a discrete time form.
\begin{equation}\label{SystemEq}
\frac{v_{t + 1}\left( \mathbf{r} \right) - v_t(\mathbf{r})}{T_s} = \int_\Omega  {k\left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{v_t}\left( \bf{r'} \right)} \right)d{\bf{r'}}}  - \alpha {v_t}\left( {\bf{r}} \right) + e_t\left( {\mathbf{r}} \right),
\end{equation}
\begin{equation}\label{SystemEq}
v_{t + 1}\left( \mathbf{r} \right)  = T_s\int_\Omega  {k\left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{v_t}\left( \bf{r'} \right)} \right)d{\bf{r'}}} - T_s\alpha {v_t}\left( {\bf{r}} \right) + v_t(\mathbf{r}) + T_se_t\left( {\mathbf{r}} \right),
\end{equation}
\begin{equation}\label{SystemEq}
v_{t + 1}\left( \mathbf{r} \right)  = T_s\int_\Omega  {k\left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{v_t}\left( \mathbf{r'} \right)} \right)d{\bf{r'}}} + (1 - T_s\alpha) v_t\left( {\bf{r}} \right) + T_se_t\left( {\mathbf{r}} \right),
\end{equation}
\begin{equation}\label{SystemEq}
v_{t + 1}\left( \mathbf{r} \right)  = T_s\int_\Omega  {k\left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{v_t}\left( \bf{r'} \right)} \right)d{\bf{r'}}} + \lambda v_t\left( {\bf{r}} \right) + T_se_t\left( {\mathbf{r}} \right).
\end{equation}
where $e_t(r)\sim \mathcal{GP}(0,R)$, $\lambda = 1-T_s\alpha$ and $T_s$ is the time step or sampling step.

@@

In order to implement standard estimation techniques we propose a decomposition of the The field using a set of Gaussian basis functions. Decomposition allows a continuous field to be represented by a finite dimensional state vector. This allows for the application of standard nonlinear, non-Gaussian state estimation methods such as sequential Monte-Carlo techniques like particle filtering. The field decomposition is described by
\begin{equation}\label{DefFieldDecomp}
{v_t}\left( {\bf{r}} \right) = {\phi ^T}\left( {\bf{r}} \right){x_t}.
\end{equation}
where $\phi(r)$ is the Gaussian basis functions that are scaled by the the state vector. The connectivity kernel can all be decomposed into a set of basis functions
\begin{equation}\label{DefKernelDecomp}
k( \mathbf{r} , \mathbf{r'} ) = \theta^T\psi(\mathbf{r} , \mathbf{r'} )
\end{equation}
Will will assume we know the parametric form of the basis functions, where the parameter $\theta$ is unknown. Each connectivity basis function can be considered a layer in the WC model, representing short range excitation, surround inhibition and long range excitation. Making substitutions we get
\begin{equation}\label{DecompModel}
{\phi ^T}\left( \mathbf{r} \right){x_{t+1}} = T_s\theta^T\int_\Omega  {\psi \left( {{\bf{r}} , {\mathbf{r'}}} \right)f\left( {{\phi ^T}\left( {{\mathbf{r'}}} \right){x_t}} \right)d{\mathbf{r'}}}  - \lambda\phi^T(\mathbf{r})x_t + T_s{e_t}\left( {\mathbf{r}} \right).
\end{equation}
To isolate the state we cross multiply by $\phi(\mathbf{r})$ and integrate over $\mathbf{r}$
\begin{equation}\label{Poo}
\int_\Omega  {\phi \left( {\bf{r}} \right){\phi ^T}\left( {\bf{r}} \right)d{\bf{r}}{x_{t+1}}}  = T_s\int_\Omega  {\phi ( \mathbf{r} )\theta^T\int_\Omega  {\psi \left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{\phi ^T}\left( {{\bf{r'}}} \right){x_t}} \right)d{\bf{r'}}} d{\bf{r}}}  - \lambda\int_\Omega \phi(\mathbf{r})\phi^T(\mathbf{r})d\mathbf{r} x_t  + T_s\int_\Omega  {\phi \left( {\mathbf{r}} \right)e_t\left( {\mathbf{r}} \right)d{\mathbf{r}}}.
\end{equation}
Now we define
\begin{equation}\label{DefGamma}
\Gamma  = \int_\Omega  {\phi \left( {\bf{r}} \right){\phi ^T}\left( {\bf{r}} \right)d{\bf{r}}}.
\end{equation}
Substituting equation~\ref{DefGamma} into \ref{Poo} and rearranging gives
\begin{equation}\label{SS Model}
{x_{t+1}} = T_s{\Gamma ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right)\theta^T\int_\Omega  {\psi \left( {{\bf{r}} , {\bf{r'}}} \right)f\left( {{\phi ^T}\left( {{\bf{r'}}} \right){x_t}} \right)d{\bf{r'}}} d{\bf{r}}} - \lambda x_t + T_s{\Gamma ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right)e_t\left( {\bf{r}} \right)d{\bf{r}}}.
\end{equation}
We can rearrange to get
\begin{equation}\label{Homogeneous SS Model}
	x_{t+1} = T_s\Gamma ^{ - 1}\int_\Omega  {\phi \left( {\bf{r}} \right)\int_\Omega  {f\left( {{\phi ^T}\left( {{\bf{r'}}} \right){x_t}} \right)\psi^T \left( {{\bf{r}} - {\bf{r'}}} \right)d{\bf{r'}}} d{\bf{r}}}\theta - \lambda x_t + T_s{\Gamma ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right)e_t\left( {\bf{r}} \right)d{\bf{r}}}.
\end{equation}
The disturbance term can be rewritten as 
\begin{equation}
	\Gamma ^{ - 1}\int_\Omega  {\phi \left( \bf{r} \right)e_t\left( \bf{r} \right)d\bf{r}} = \int_\Omega  {\phi \left( \bf{r} \right)\tilde{e}_t\left( \bf{r} \right)d\bf{r}},
\end{equation}
where $\tilde{e}_t(r)~N(0,R'R)$ and $R'$ is the Cholesky decomposition of $\Gamma^{-1}$. This provides the final state-space model
\begin{equation}\label{AbbrevSSModel}
x_{t + 1} = T_sq(x_t)\theta -\lambda x_t + T_s\int_\Omega  {\phi ( \mathbf{r} )\tilde{e}_t( \mathbf{r} )d\mathbf{r}}
\end{equation}
with observation equation
\begin{equation}\label{ObservationEquation}
    y_t( k ) = \int_\Omega  m( k,\mathbf{r} )\phi^T( \mathbf{r} ) d\mathbf{r}x_t  + \varepsilon ( k ),
\end{equation}

\subsection{Discrete Space For Simulation}

In this section we describe a discrete time, discrete space model that we are using for the simulations. Now we define the spatial aspect of the model on a regular square $i,j$ grid of neural masses, where the spatial step size $\Delta \mathbf{r}_i = \Delta \mathbf{r}_j = \Delta \mathbf{r}$ giving
\begin{equation}\label{DiscreteSpaceModelij}
	x_{t+1} = T_s\Gamma ^{ - 1}\sum_i \sum_j  {\phi \left( \mathbf{r} \right)\sum_i\sum_j  {f\left( \phi^T \left( \mathbf{r'} \right) x_t \right)\psi^T \left( \mathbf{r} - \bf{r'} \right)\Delta\mathbf{r'}^2} \Delta\mathbf{r}^2} \theta - \lambda x_t + T_s\sum_i\sum_j  {\phi \left( {\bf{r}} \right)\tilde{e}_t\left( \bf{r} \right)\Delta\mathbf{r}^2}.
\end{equation}
\section{Estimation of the Nonlinear Homogeneous IDE Neural Field Model}
